### 2021-05-21 paper/main.tex
Finish first paper draft

\documentclass[sigplan,screen]{acmart}

\usepackage{xcolor}
\definecolor{codegreen}{rgb}{0,0.6,0}

\usepackage{cancel}

\usepackage{listings}
\lstset{%
    language=Haskell,
    columns=fullflexible,
    basicstyle=\ttfamily,
    commentstyle=\color{codegreen}
}

% shorthand for big O
\renewcommand\O[1]{$\mathcal{O}(#1)$}

%% Rights management information.  This information is sent to you
%% when you complete the rights form.  These commands have SAMPLE
%% values in them; it is your responsibility as an author to replace
%% the commands and values with those provided to you when you
%% complete the rights form.
\setcopyright{acmlicensed}
\copyrightyear{2021}
\acmYear{2021}
\acmDOI{tbd}

%% These commands are for a PROCEEDINGS abstract or paper.
\acmConference[Haskell '21]{the 14th ACM SIGPLAN International Haskell Symposium}{August 26--27, 2021}{Virtual Event}
%\acmBooktitle{Woodstock '18: ACM Symposium on Neural Gaze Detection,
%  June 03--05, 2018, Woodstock, NY}
%\acmPrice{15.00}
%\acmISBN{978-1-4503-XXXX-X/18/06}

%%
%% Submission ID.
%% Use this when submitting an article to a sponsored event. You'll
%% receive a unique submission ID from the organizers
%% of the event, and this ID should be used as the parameter to this command.
%%\acmSubmissionID{123-A56-BU3}

%%
%% end of the preamble, start of the body of the document source.
\begin{document}

%%
%% The "title" command has an optional parameter,
%% allowing the author to define a "short title" to be used in page headers.
\title{Liquid Amortization - Proving amortized complexity with LiquidHaskell (Functional Pearl)}

%%
%% The "author" command and its associated commands are used to define
%% the authors and their affiliations.
%% Of note is the shared affiliation of the first two authors, and the
%% "authornote" and "authornotemark" commands
%% used to denote shared contribution to the research.
\author{Jan van Brügge}
\email{jan.van-bruegge@tum.de}
%\orcid{1234-5678-9012}
\affiliation{%
  \institution{Technical University Munich}
  \streetaddress{Arcisstraße 21}
  \city{Munich}
  \state{Bavaria}
  \country{Germany}
  \postcode{80333}
}

%%
%% By default, the full list of authors will be used in the page
%% headers. Often, this list is too long, and will overlap
%% other information printed in the page headers. This command allows
%% the author to define a more concise list
%% of authors' names for this purpose.
%\renewcommand{\shortauthors}{Trovato and Tobin, et al.}

%%
%% The abstract is a short summary of the work to be presented in the
%% article.
\begin{abstract}
Proofs of complexity of production code is usually done in external theorem provers. This makes it necessary to translate the data structures to be proved from the production language into the language of the prover. Such a translation step may introduce errors by a mismatch in features in the two languages. We show how to prove amortized complexity of data structures directly in Haskell using LiquidHaskell~\cite{liquidhaskell}. Besides skipping the translation step, our approach also has a didactic advantage as new learners do not need to learn a new language for their proofs and can focus on the new concepts only. We do not assume prior knowledge of amortized complexity as we explain the concepts and apply them in our first case study, a simple stack with multipop. The usefulness of our approach in a production setting is verified by the second case study, the finger tree which powers \texttt{Data.Seq} from the \textit{containers} package which is widely depended on. Finally we discuss the current limitations of LiquidHaskell in this area and how they might be lifted.
\end{abstract}

%%
%% The code below is generated by the tool at http://dl.acm.org/ccs.cfm.
%% Please copy and paste the code instead of the example below.
%%
\begin{CCSXML}
<ccs2012>
   <concept>
       <concept_id>10003752.10003809.10010031</concept_id>
       <concept_desc>Theory of computation~Data structures design and analysis</concept_desc>
       <concept_significance>500</concept_significance>
       </concept>
   <concept>
       <concept_id>10011007.10011006.10011008.10011009.10011012</concept_id>
       <concept_desc>Software and its engineering~Functional languages</concept_desc>
       <concept_significance>500</concept_significance>
       </concept>
   <concept>
       <concept_id>10011007.10010940.10010992.10010998.10010999</concept_id>
       <concept_desc>Software and its engineering~Software verification</concept_desc>
       <concept_significance>500</concept_significance>
       </concept>
 </ccs2012>
\end{CCSXML}

\ccsdesc[500]{Theory of computation~Data structures design and analysis}
\ccsdesc[500]{Software and its engineering~Functional languages}
\ccsdesc[500]{Software and its engineering~Software verification}

%%
%% Keywords. The author(s) should pick words that accurately describe
%% the work being presented. Separate the keywords with commas.
\keywords{datastructures, LiquidHaskell, amortized complexity, finger trees, theorem proving, functional pearl}

%%
%% This command processes the author and affiliation and title
%% information and builds the first part of the formatted document.
\maketitle

\section{Introduction}

Formally proving certain properties of production code is getting more and more common. For either safety critical code that is not allowed to "go wrong" or code that is fundamental and reused very often, bugs can have a widespread impact. Proving that the code behaves according to the specification can give an extra layer of confidence in such cases. Fundamental data structures are one of the most important parts of a language ecosystem and a lot of production code depends on them. In the Haskell ecosystem, this job is done by the \textit{containers} package~\cite{containers} which has over 5300 reverse dependencies. A lot of code is depending on the assumptions that the data structures perform well and adhere to the stated computational complexities from the documentation. Thus, proving those assumptions correct benefits a huge number of people.

Usually these proofs are done in an external theorem prover, requiring an explicit translation step from the production language to the language of the theorem prover. This translation step has a few disadvantages. First, the semantics of the target language may differ from the semantics of the production language. This is especially true for Haskell with its pervasive laziness, compared with most theorem provers that operate with a strict language. Second, such a translation step may introduce subtle differences or bugs, either by accident in the manual case or through a bug in the automated case. Lastly, the target language may not support certain features present in the source language such as polymorphic recursion in data types. In those cases, the data structure needs to be encoded in the target language, again opening the doors for small and subtle errors. Because the code that is actually used to conduct the proofs might me far removed from the source code that is running in production, it is often hard to reconnect the proofs back to the source code.

A different approach comes in the form of LiquidHaskell~\cite{liquidhaskell}. It is a refinement type~\cite{refinement_types} checker for Haskell that since 2020 is implemented as a type checker plugin for GHC, the main Haskell compiler used in the industry. It allows to specify additional constraints on types like "integers greater than zero" instead of just "integers". These constraints are propagated through the code and checked by an external SMT prover. With a handful of proof combinators defined by LiquidHaskell, it is possible to use Haskell directly as a theorem prover, enabling intuitive equational reasoning within the language~\cite{tpfa}. This avoids any translation step and makes it easy to keep proof and implementation in sync, as any failure to do so will result in an immediate compiler error. Another benefit is that the reader or writer of a proof does not need to learn a new language, the production language is the proof language.

This paper thus:
\begin{itemize}
\item{Assumes no prior knowledge of amortized complexity by recapping the theoretical background}
\item{Shows the reasoning infrastructure that is needed to prove amortized complexity with LiquidHaskell}
\item{Applies this technique to the finger tree, the data structure behind \texttt{Data.Seq} from \textit{containers}, as a case study}
\item{Discusses the usefulness of LiquidHaskell for such proofs compared with other theorem provers}
\end{itemize}

\section{Amortized complexity}\label{sec:complexity}

It is common to check how many time units an operation on a data structure takes in relation to the number of elements in the data structure. As the performance can vary wildly for the same operation depending on the state of data structure --- a linear search takes $1$ time unit if the searched element is at the front, while it takes $n$ time units if it is the last element --- usually one uses the worst case for an operation as a metric and uses Landau notation to specify it. With our linear search example we would say that it is in \O{n} because it will not perform worse than $n$ time units.

But often this worst case is too pessimistic. When an operation changes the data structure, for example by removing elements, consecutive uses of operation may be cheaper. Or to use an expensive operation it might be necessary to have a certain amount of cheap operations. In this case one uses \textit{amortized complexity} which does not only look at a single operation in isolation, but on chains of operations. In particular, it assumes that that the intermediate data structures are not shared, i.e. they are not persisted. There are multiple methods to proof amortized complexity.

As a running example we will use a stack with \textit{push} and \textit{multipop} operations. The former allocates a new element and a references to the rest of the stack and thus taking a constant amount of time. The latter removes the first $k$ elements from the stack. In the worst case, we remove all the elements, making \texttt{multipop} \O{n}. One can already see that this is too pessimistic, as after removing all elements from the stack, the next \texttt{multipop} will do nothing. In fact, every element we remove needs to be added to the stack at some point, so we can spread out the cost of \texttt{multipop} over many pushes. There are multiple techniques to prove amortized complexity, we present the two most popular ones.

\subsection{Banker's method}\label{sec:banker}

The banker's method works by having an bank account that can store time units. Operations have to pay their actual cost and may additionally pay extra into the account. Expensive operations can then use the stored time units in the bank to pay for themselves. In our example a \texttt{push} has to pay one actual time unit and it will also pay one extra time unit into the bank account. This means that the bank account always contains as many time units as there are elements on the stack. That way, any call to \texttt{multipop} can use the bank account to pay for itself and the account will never become negative.

\subsection{Physicist's method}\label{sec:physicist}

The banker's method is hard to use for a formal proof because the bank account acts as state between the operations. A simpler method is the physicist's method that does not need such an account. The idea is to define a potential $\Phi$ ("phi") for the data structure. This potential should be $0$ for an empty data structure and $\ge 0$ for all other states. Intuitively the potential will declare how much time needs to be saved up to pay for expensive operations. We then define the amortized cost of an operation as $c + \Phi(h') - \Phi(h)$ where $c$ is the actual cost for the operation, $h$ is the data structure before the operation and $h'$ is the data structure afterwards. Looking at a chain of operations $1, 2, ..., n$, the sum of this amortized time is an overestimation of the actual cost by $\Phi(h_n)$ --- because $\Phi(h_0)$ is by definition $0$ --- and thus an upper bound on the actual cost (see equation~\ref{eq:sum}).

\begin{gather}
\nonumber (c_1 + \cancel{\Phi(h_1)} - \Phi(h_0)) + (c_2 + \cancel{\Phi(h_2)} - \cancel{\Phi(h_1)}) \\
\nonumber    + ... + (c_n + \Phi(h_n) - \cancel{\Phi(h_{n - 1})}) = \\
  c_1 + c_2 + ... + c_n + \Phi(h_n) - \Phi(h_0) \label{eq:sum}
\end{gather}

For our stack we can define the potential to be the height of the stack. This fulfils the requirements as the empty stack has a potential of zero and all other states have a potential greater or equal to zero. For the actual proof of this example see the first case study in section~\ref{sec:stack}.

\section{LiquidHaskell for theorem proving}

As described by Vazou et. al.~\cite{tpfa}, one can use LiquidHaskell to do complex equational reasoning that is machine checked by the compiler. It uses refinement types~\cite{refinement_types} that add additional constraints on top of normal Haskell data types which are checked by an external SMT solver. By default LiquidHaskell uses the z3~\cite{z3} solver, but it also supports cvc4~\cite{cvc4}. For example we can declare the type of integers greater or equal to zero with a type synonym. LiquidHaskell annotions use comments with an @ sign, so that the code still compiles even without LiquidHaskell. It also allows to name arguments, so they can be referred to later in the refinement. At the moment to define functions with refinement, one has to copy the Haskell type signature into a LiquidHaskell comment and add the refinements there. A future version that is more deeply integrated into the Haskell compiler may not need this code duplication.

\begin{lstlisting}
{-@ type Nat = { x:Int | x >= 0 } @-}
\end{lstlisting}

In LiquidHaskell, all functions need to terminate. Usually a suitable termination metric is automatically guessed, but in some more complicated cases, we can also manually supply this metric (see section~\ref{sec:append} for an example of this). A simple recursive function can be marked as \textit{measure} which allows the use of the function in those termination metrics as well as in the refinement itself. With the termination, LiquidHaskell is able to generate an induction principle for the function that can be used in a proof. Because LiquidHaskell is just Haskell, this induction amounts to a recursive call to the proof and is also used when defining ordinary Haskell functions with refinement on their types. For example the fact that the length of a list is never negative.

\begin{lstlisting}
{-@ measure length @-}
{-@ length :: [a] -> Nat @-}
length :: [a] -> Int
length [] = 0 -- trivial, 0 >= 0
length (x:xs) = 1 + length xs
    -- by induction: length xs >= 0
    -- so 1 + length xs >= 0
\end{lstlisting}

LiquidHaskell also provides a set of proof combinators and a \texttt{Proof} type. This type is just a type alias for the normal Haskell unit type. The reasoning comes from the \texttt{===} operator that asserts that both sides of the equality are truly equal. As a simple example we can proof that the length of an \texttt{append} function is the sum of the lengths of the two original lists. As append is not a \textit{measure} (it does not return a natural number), we have to use \textit{reflection}~\cite{reflection} to lift it to the logic level of LiquidHaskell. The proof itself is done by case distinction on the first argument. This distinction is done by pattern matching on it like one would in a normal Haskell function. We start with stating the left hand side of the equation how it is in the refinement type and try to get to the right hand side of the equality by simplification and induction.

\begin{lstlisting}
{-@ reflect append @-}
append :: [a] -> [a] -> [a]
append [] ys = ys
append (x:xs) ys = x:(append xs ys)

{-@ lengthP :: xs:[a] -> ys:[a] ->
  { length (append xs ys) ==
      length xs + length ys }
@-}
lengthP :: [a] -> [a] -> Proof

-- base case
lengthP [] ys = length (append [] ys)
  -- Use definition of append [] _
    === length ys
    === 0 + length ys
  -- use definition of length []
    === length [] + length ys
    *** QED

-- recursive case
lengthP (x:xs) ys =
    length (append (x:xs) ys)
  -- Use definition of append
    === length (x:(append xs ys))
  -- Use definition of length
    === 1 + length (append xs ys)
  -- Use induction with the ? operator
    ? lengthP xs ys
    === 1 + length xs + length ys
  -- Use definition of length
    === length (x:xs) + length ys
    *** QED
\end{lstlisting}

The triple star operator (\texttt{***}) and the \texttt{QED} type at the end are just to cast our proof to Haskell's unit type and to make the proof look nicer.

\section{Case studies}

\subsection{Stack with Multipop}\label{sec:stack}

To continue with the example from section~\ref{sec:complexity}, we want to prove that $multipop$ has an amortized complexity of \O{1}. This data structure and its operations could be defined like this in Haskell. We already add some refinements to assert that $k$ is not negative.

\begin{lstlisting}
data Stack a = Empty | Elem a (Stack a)

push :: a -> Stack a -> Stack a
push = Elem

{-@ multipop :: Nat -> Stack a -> Stack a @-}
multipop :: Int -> Stack a -> Stack a
multipop _ Empty = Empty
multipop 0 s = s
multipop n (Elem _ s) = multipop (n - 1) s
\end{lstlisting}

To formally prove this with the physicist's method, we first need to define a potential $\Phi$ and a timing function for each of our operations that gives us the actual cost of that operation. As seen in section~\ref{sec:physicist}, we use the height of the stack as our potential. Thus we need to define a function that calculates the height of a stack. To define the timing functions, we copy the code for the operation, replace constant time steps with $1$ and adjust the recursion.

\begin{lstlisting}
{-@ phi :: Stack a -> Nat @-}
phi :: Stack a -> Int
phi Empty = 0
phi (Elem _ s) = 1 + phi s

{-@ pushT :: a -> Stack a -> Nat @-}
pushT :: a -> Stack a -> Int
pushT _ _ = 1

{-@ multipopT :: Nat -> Stack a -> Nat @-}
multipopT :: Int -> Stack a -> Int
multipopT _ Empty = 1
multipopT 0 _ = 1
multipopT n (Elem _ s) = 1 + multipopT (n - 1) s
\end{lstlisting}

To use those functions in a LiquidHaskell proof, we also need to reflect them into the logic level~\cite{reflection} by adding \texttt{reflect} annotations.

\begin{lstlisting}
{-@ reflect push @-}
{-@ reflect pushT @-}
{-@ reflect multipop @-}
{-@ reflect multipopT @-}
{-@ reflect phi @-}
\end{lstlisting}

First we will prove the complexity of \texttt{push}. While the amortized cost is the same as the actual cost, we need to make sure that our potential function does not cause those two to diverge. As LiquidHaskell does not allow quantifiers to ensure termination of the SMT solver, we need to guess an upper limit that our amortized runtime will have. As seen earlier in section~\ref{sec:banker}, we can pay two time units for a push to amortize the pops, so we will use $2$ as our guess. To specify our proof goal, we need to name the parameters in the LiquidHaskell annotation. After than we can use equational reasoning to proof our goal.

\begin{lstlisting}
import Language.Haskell.Liquid.ProofCombinators

{-@ pushP :: x:a -> s:Stack a ->
  { pushT x s + phi (push x s) - phi s <= 2 }
@-}
pushP :: a -> Stack a -> Proof
pushP x s = pushT x s + phi (push x s) - phi s
-- Use definition of push and pushT
  === 1 + phi (Elem x s) - phi s
-- Use definition of phi (Elem _ _)
  === 1 + (1 + phi s) - phi s
-- phi s gets canceled, 1 + 1 <= 2
  *** QED
\end{lstlisting}

To proof complexity of \texttt{multipop}, we will need to make a case distinction between the two base cases and the recursive case. With LiquidHaskell this is just a normal pattern match. Leaving the recursive case undefined for now, proving the base cases is very similar to \texttt{push}.

\begin{lstlisting}
{-@ multipopP :: k:Nat -> s:Stack a ->
  { multipopT k s + phi (multipop k s)
      - phi s <= 2 }
@-}
multipopP :: Int -> Stack a -> Proof
multipopP 0 s = multipopT 0 s
    + phi (multipop 0 s) - phi s
-- Use definition of multipop/T 0 _
  === 1 + phi s - phi s
-- phi s cancels, 1 <= 2
  *** QED
multipopP k Empty = multipopT k Empty
    + phi (multipop k Empty) - phi Empty
-- Use definition of multipop/T _ Empty
  === 1 + phi Empty - phi Empty
-- phi Empty cancels, 1 <= 2
  *** QED
\end{lstlisting}

To avoid cluttering the proof of the recursive case, we use an \textit{as pattern} to alias \texttt{Elem x s} to \texttt{xs}. The first steps again are applying the definitions of the used functions and simplifying the result. In the end we use the \texttt{?} operator to add the induction hypothesis to the proof step and thus complete the proof.

\begin{lstlisting}
multipopP k xs@(Elem x s) = multipopT k xs
    + phi (multipop k xs) - phi xs
-- Use definition of multipopT for recusive case
  === (1 + multipopT (k - 1) s)
    + phi (multipop k xs) - phi xs
-- Use definition of multipop for recusive case
  === (1 + multipopT (k - 1) s)
    + phi (multipop (k - 1) s) - phi xs
-- Use definition of phi (Elem _ _)
  === (1 + multipopT (k - 1) s)
    + phi (multipop (k - 1) s) - (1 + phi s)
-- Remove parenthesis
  === 1 + multipopT (k - 1) s
    + phi (multipop (k - 1) s) - 1 - phi s
-- 1s cancel out
  === multipopT (k - 1) s
    + phi (multipop (k - 1) s) - phi s
-- use induction
  ? multipopP (k - 1) s
  *** QED
\end{lstlisting}

\subsection{Finger Tree}

To show that this approach scales also to more complicated data structures that are widely used in the industry, we will prove the complexity of a finger tree, the data structure behind \texttt{Data.Seq} from \textit{containers}~\cite{containers}. It is originally described by Hinze and Paterson~\cite{fingertrees}, which the implementation from \textit{containers} closely follows. The following definition is taken directly from \textit{containers}, but slightly modified to remove the size parameter. LiquidHaskell caused problems in the proofs when using type classes, so because the functions we will use do not use the size we omitted it. We will discuss such limitations more in depth in section~\ref{sec:limitations}.

\begin{lstlisting}
data FingerTree a = Empty
  | Single a
  | Deep (Digit a) (FingerTree (Node a)) (Digit a)

data Node a = Node2 a a | Node3 a a a

data Digit a = One a
  | Two a a
  | Three a a a
  | Four a a a a
\end{lstlisting}

The finger tree has between one and four \textit{digits} on either side to enable cheap \texttt{cons} and \texttt{snoc}. As shown by Claessen~\cite{fingertrees_new}, this would also work with a maximum of just three digits, but we will go with the definition from \textit{containers}. The most interesting point about the data structure is the polymorphic recursion that makes the data structure type safe as it guarantees equal nesting of \texttt{Node} types on both sides of the spine. It is also what makes it difficult to verify in other languages, because for example the Isabelle/HOL~\cite{isabelle} and the Lean~\cite{lean} theorem provers both do not directly support polymorphic recursion. In these languages, it is necessary to define the type such that it would allow uneven nesting of \texttt{Node}s and then "carve out" only the valid finger trees. Other provers like Coq~\cite{coq} support polymorphic recursion making a more direct one to one translation possible.

\subsubsection{Cons and Snoc}

As the finger tree is a symmetric tree, \texttt{cons} and \texttt{snoc} are very similar. For this reason we will show an explicit proof only for \texttt{cons} and provide the proof for \texttt{snoc} in the appendix. First we again copy definition of \texttt{cons} from the \textit{containers} package, adjusting it slightly to remove the size information that would require the type class. Also, to later use the function in a proof we add a \texttt{reflect} pragma.

\begin{lstlisting}
{-@ reflect cons @-}
cons :: a -> FingerTree a -> FingerTree a
cons a EmptyT = Single a
cons a (Single b) =
  Deep (One a) EmptyT (One b)
cons a (Deep (One b) m sf) =
  Deep (Two a b) m sf
cons a (Deep (Two b c) m sf) =
  Deep (Three a b c) m sf
cons a (Deep (Three b c d) m sf) =
  Deep (Four a b c d) m sf
cons a (Deep (Four b c d e) m sf) =
  Deep (Two a b) (cons (Node3 c d e) m) sf
\end{lstlisting}

The only complicated case is if we have already four digits at the front, as in all other cases we can simply increase the amount by one. In the case that we can not do so, we put the new data in front and push the "overflow" in to the recursion. The next step is to think of a valid potential function $\Phi$. Again, the intuition is which state of the data structure makes \texttt{cons} expensive. As seen above it is when the digit in the front is already full. So we can define the \textit{danger} of a digit as every state that will make the operation expensive. To also support and proof \texttt{uncons}, the inverse operation, we not only mark a full digit as dangerous, but also one that has only one element inside (as this would make \texttt{uncons} go into the recursion). Our potential is then the sum of the danger in the tree.

\begin{lstlisting}
{-@ reflect dang @-}
{-@ dang :: Digit a -> Nat @-}
dang :: Digit a -> Int
dang One{} = 1
dang Two{} = 0
dang Three{} = 0
dang Four{} = 1

{-@ reflect phi @-}
{-@ phi :: FingerTree a -> Nat @-}
phi :: FingerTree a -> Int
phi EmptyT = 0
phi Single{} = 0
phi (Deep pr m sf) =
  dang pr + phi m + dang sf
\end{lstlisting}

The timing function for cons is similar to the timing functions in the stack just a copy of the original function with all non-recursive cases return $1$.

\begin{lstlisting}
{-@ reflect consT @-}
{-@ consT :: a -> FingerTree a -> Nat @-}
consT :: a -> FingerTree a -> Int
consT _ EmptyT = 1
consT _ Single{} = 1
consT _ (Deep One{} _ _) = 1
consT _ (Deep Two{} _ _) = 1
consT _ (Deep Three{} _ _) = 1
consT _ (Deep (Four _ c d e) m _) =
  1 + consT (Node3 c d e) m
\end{lstlisting}

With all the preliminaries out of the way we can start proving the amortized complexity. Like with the stack we need to guess a constant that is an upper bound on our amortized run time. As we do not have a simple mental model like with the stack, we will delay this choice until we have proven the base cases. We will also use \textit{as patterns} again to avoid repetition.

\begin{lstlisting}
{-@ consP :: x:a -> t:FingerTree a ->
  { consT x t + phi (cons x t) - phi t <= c }
@-}
consP :: a -> FingerTree a -> Proof
consP x t@EmptyT =
    consT x t + phi (cons x t) - phi t
-- Use definition of cons/T _ EmptyT
  === 1 + phi (Single x) - phi EmptyT
-- Use definition of phi
  === 1 + 0 - 0
-- 1 must be <= c, so c >= 1
  *** QED
consP x t@(Single a) =
    consT x t + phi (cons x t) - phi t
-- Use definition of cons/T _ (Single a)
  === 1 + phi (Deep (One x) EmptyT (One a))
        - phi (Single a)
-- Use definition of phi
  === 1 + dang (One x) + phi EmptyT
        + dang (One a) - 0
-- Use definition of phi and dang
  === 1 + 1 + 0 + 1 - 0
-- 3 must be <= c, so c >= 3
  *** QED
\end{lstlisting}

We omit the proof cases for one to three digits, as they work exactly as the other two base cases. The constant however will not change in the other base cases, so $3$ really is our upper bound to use here. The interesting case is when we have four digits. In this case we need to use induction to proof our goal.

\begin{lstlisting}
consP x t@(Deep (Four a b c d) m sf) =
    consT x t + phi (cons x t) - phi t
-- Use definition of consT
  === (1 + consT (Node3 b c d) m)
  + phi (cons x t) - phi t
-- Use definition of cons
  === (1 + consT (Node3 b c d) m)
  + phi (Deep (Two x a) (cons (Node3 b c d) m) sf)
  - phi t
-- Use definition of phi
  === (1 + consT (Node3 b c d) m)
  + dang (Two x a) + phi (cons (Node3 b c d) m)
  + dang sf
  - (dang (Four a b c d) + phi m + dang sf)
-- Remove parenthesis
  === 1 + consT (Node3 b c d) m
  + dang (Two x a) + phi (cons (Node3 b c d) m)
  + dang sf
  - dang (Four a b c d) - phi m - dang sf
-- dang sf cancels, use definition of dang
  === 1 + consT (Node3 b c d) m
  + 0 + phi (cons (Node3 b c d) m) - 1 - phi m
-- 1s cancel
  === consT (Node3 b c d) m
  + phi (cons (Node3 b c d) m) - phi m
-- Use induction
  ? consP (Node3 b c d) m
  *** QED
\end{lstlisting}

\subsubsection{Append}\label{sec:append}

Similar to \texttt{push} in the stack, even though \texttt{append} has the same amortized complexity as its normal complexity, we need to show that it never increases the potential more than a logarithmic amount. Contrary to the simplified definition of Claessen~\cite{fingertrees_new}, the \textit{containers} package uses two sets of code generated, mutually recursive functions to implement \texttt{append}. One set of functions called \texttt{appendTreeX} handles the base cases where one of the two trees is either empty or consists of a singular element by using \texttt{cons} and \texttt{snoc}. In the recursive case it calls the \texttt{addDigitsX} functions that group arguments as \texttt{Node}s and itself call the \texttt{appendTreeX} functions. Because the \texttt{addDigitsX} functions do pattern match on their tree arguments at all, LiquidHaskell can not find a termination metric by itself. For this, we need to track the arguments across functions calls because \texttt{appendTreeX} has a strictly decreasing argument. We can do so by adding a ghost parameter to order the functions for LiquidHaskell. This order is lower for all \texttt{appendTreeX} functions than the \texttt{addDigitsX} functions. Additionally we need to set the case expansion to a high number such that all pattern matches are evaluation, else the termination proof fails. See the following snippet as an example with most cases and function omitted for brevity.

\begin{lstlisting}
{-@ LIQUID --max-case-expand=1000 @-}

{-@ appendTree1 :: a:FingerTree (Node a) ->
  Node a -> FingerTree (Node a) ->
  FingerTree (Node a) / [size a, 1] @-}
{-@ reflect appendTree1 @-}
appendTree1 :: FingerTree (Node a) -> Node a ->
    FingerTree (Node a) -> FingerTree (Node a)
appendTree1 EmptyT !a xs = a `consTree` xs
appendTree1 xs !a EmptyT = xs `snocTree` a
appendTree1 (Single x) !a xs =
    x `consTree` a `consTree` xs
appendTree1 xs !a (Single x) =
    xs `snocTree` a `snocTree` x
appendTree1 (Deep pr1 m1 sf1) a (Deep pr2 m2 sf2) =
    Deep pr1 (addDigits1 m1 sf1 a pr2 m2) sf2

{-@ addDigits1 :: a:FingerTree (Node (Node a)) ->
  Digit (Node a) -> Node a -> Digit (Node a) ->
  FingerTree (Node (Node a)) ->
  FingerTree (Node (Node a)) / [size a, 5] @-}
{-@ reflect addDigits1 @-}
addDigits1 :: FingerTree (Node (Node a)) ->
  Digit (Node a) -> Node a -> Digit (Node a) ->
  FingerTree (Node (Node a)) ->
  FingerTree (Node (Node a))
addDigits1 m1 (One a) b (One c) m2 =
  appendTree1 m1 (Node3 a b c) m2
addDigits1 m1 (One a) b (Two c d) m2 =
  appendTree2 m1 (Node2 a b) (Node2 c d) m2
-- further cases omitted
\end{lstlisting}

This case expansion results in an explosion of constraints for the SMT solver to check. While the whole proof for \texttt{cons} including the defined functions generates less than 100 constraints, this termination proof generates over 3000. This leads to extremely high compile times. This is made worse by the need to define a timing function that needs to follow this mutually recursive structure and generates an additional 8000 constraints such that loading the file takes several minutes. Another problem is the logarithmic runtime itself. SMT provers are great with linear arithmetic, but are significantly slower with multiplication. Division is usually left as an unevaluated function. While one can transform the proof goal to turn logarithms into powers, it still requires assuming basic arithmetic rules on the logic level.

For those reasons, using LiquidHaskell to proof the complexity with the physicist's method is not feasible (see section~\ref{sec:limitations} for a discussion of those limitations). Claessen~\cite{fingertrees_new} also does not formally proof this in his functional pearl, but provides a different reasoning to why \texttt{append} still is logarithmic. The potential only depends on the spine of the tree, so a new layer in the tree only adds at most a constant amount to the potential ($2$ if both digits are dangerous) making it grow linearly. Because the nodes in the digits always at least double (in the case of two \texttt{Node2}) the overall size of the tree grows exponentially. Thus the potential grows logarithmic compared to the size of the tree.

\subsection{Definition of the timing function}

One of the points where the analysis may lead to wrong results is if the timing function does not directly mirror the original function. So when the implementation changes and the timing function does not, it might be possible for the proofs to still pass. While this is unlikely, given how verbose proofs with LiquidHaskell are, it is not negligible.

If one is in control of the source code and can change the implementation, it would be possible to factor out the recursion and share it between the implementation and the timing function. This however would clutter the implementation and most likely decrease performance as it is often the case with recursion schemes.

A better solution would be to derive the timing function using meta programming in the form of TemplateHaskell. This meta function would take the name of the function in the implementation and create a new one with a configurable suffix. To do so it would check if the right hand sides contain a recursive call to the given name. If not, it replaces the right hand side with $1$, otherwise it changes the recursion to use the name with suffix and add $1$ at the end. The resulting function could then be spliced back into the source code.

\begin{lstlisting}
cons :: a -> FingerTree a -> FingerTree a
cons a EmptyT = Single a
cons a (Deep (Four b c d e) m sf) =
    Deep (Two a b) (cons (Node3 c d e) m) sf
-- other cases omitted

$(mkTimingFn "T" ''cons)
-- generates this; variable names omitted:
-- consT :: a -> FingerTree a -> Int
-- consT _ EmptyT = 1
-- consT _ (Deep (Four _ c d e) m _) =
--    consT (Node3 c d e) m + 1
-- other cases omitted
\end{lstlisting}

\subsection{LiquidHaskell for verification}

As shown in earlier sections, LiquidHaskell has some limitations at the moment with regards to proving more complex functions. First and foremost the support for type classes or at least the support for instances involving Haskell \texttt{newtype}s is not great. Currently, LiquidHaskell only supports measure classes, which need to be defined by recursion over a single data type similar to a normal measure. For this reason it was not possible to use it for the \texttt{Sized} type class from \textit{containers}. There are several open issues about the type class support on the LiquidHaskell bug tracker, so we expect this to get better in the future.

Another pain point was the combinatorial explosion when generating the constraints for the termination of \texttt{append} (see section~\ref{sec:append}). It might be possible to be smarter when generating the constraints here and thus creating less work for the SMT solver. LiquidHaskell also provides an option to disable the termination checker, but enabling this option surprisingly did not improve compile time. In fact, several other options that are supposed to improve performance lik \texttt{{-}{-}diff} which enables incremental type checking or \texttt{{-}{-}fast} which disables inference of polymorphic refinement, did either not affect performance at all or (in the case of \texttt{{-}{-}fast}) made it even worse.

\section{Conclusion}

Despite the current limitations of LiquidHaskell preventing more complicated functions from being proved formally, there is still a lot of value in this approach. Proving complexity of core data structures immediately pays back in a huge gain in confidence across the many applications that depend on them. While at the moment, amortized complexity of \O{1} seems to be the most feasable to proof, we expect the tooling around LiquidHaskell to improve a lot in the future, be it by incremental changes like better support for type classes or by more complicated means like a deeper embedding into a future dependently typed Haskell core.

Aside from the industry application, we think this approach also has a big didactic advantage. To learn more about the complexity of data structures, a learner does not need to learn a new language on top of new concepts, but instead can use the language they already know to apply their new knowledge. The recent improvements in the usability of LiquidHaskell --- turning the previously standalone executable into a GHC plugin~\cite{lh_plugin} --- play a big role here as it drastically reduces the barrier of entry.

%%
%% The acknowledgments section is defined using the "acks" environment
%% (and NOT an unnumbered section). This ensures the proper
%% identification of the section in the article metadata, and the
%% consistent spelling of the heading.
%\begin{acks}
%\end{acks}

%%
%% The next two lines define the bibliography style to be used, and
%% the bibliography file.
\bibliographystyle{ACM-Reference-Format}
\bibliography{references}

%%
%% If your work has an appendix, this is the place to put it.
\appendix

\section{Full proof code for the stack}

\begin{lstlisting}
import Language.Haskell.Liquid.ProofCombinators

data Stack a = Empty | Elem a (Stack a)

{-@ reflect push @-}
push :: a -> Stack a -> Stack a
push = Elem

{-@ multipop :: Nat -> Stack a -> Stack a @-}
multipop :: Int -> Stack a -> Stack a
multipop _ Empty = Empty
multipop 0 s = s
multipop n (Elem _ s) = multipop (n - 1) s

{-@ reflect phi @-}
{-@ phi :: Stack a -> Nat @-}
phi :: Stack a -> Int
phi Empty = 0
phi (Elem _ s) = 1 + phi s

{-@ reflect pushT @-}
{-@ pushT :: a -> Stack a -> Nat @-}
pushT :: a -> Stack a -> Int
pushT _ _ = 1

{-@ reflect multipopT @-}
{-@ multipopT :: Nat -> Stack a -> Nat @-}
multipopT :: Int -> Stack a -> Int
multipopT _ Empty = 1
multipopT 0 _ = 1
multipopT n (Elem _ s) = 1 + multipopT (n - 1) s

{-@ pushP :: x:a -> s:Stack a ->
  { pushT x s + phi (push x s) - phi s <= 2 }
@-}
pushP :: a -> Stack a -> Proof
pushP x s = pushT x s + phi (push x s) - phi s
-- Use definition of push and pushT
  === 1 + phi (Elem x s) - phi s
-- Use definition of phi (Elem _ _)
  === 1 + (1 + phi s) - phi s
-- phi s gets canceled, 1 + 1 <= 2
  *** QED

{-@ multipopP :: k:Nat -> s:Stack a ->
  { multipopT k s + phi (multipop k s)
      - phi s <= 2 }
@-}
multipopP :: Int -> Stack a -> Proof
multipopP 0 s = multipopT 0 s
    + phi (multipop 0 s) - phi s
-- Use definition of multipop/T 0 _
  === 1 + phi s - phi s
-- phi s cancels, 1 <= 2
  *** QED
multipopP k Empty = multipopT k Empty
    + phi (multipop k Empty) - phi Empty
-- Use definition of multipop/T _ Empty
  === 1 + phi Empty - phi Empty
-- phi Empty cancels, 1 <= 2
  *** QED
multipopP k xs@(Elem x s) = multipopT k xs
    + phi (multipop k xs) - phi xs
-- Use definition of multipopT for recusive case
  === (1 + multipopT (k - 1) s)
    + phi (multipop k xs) - phi xs
-- Use definition of multipop for recusive case
  === (1 + multipopT (k - 1) s)
    + phi (multipop (k - 1) s) - phi xs
-- Use definition of phi (Elem _ _)
  === (1 + multipopT (k - 1) s)
    + phi (multipop (k - 1) s) - (1 + phi s)
-- Remove parenthesis
  === 1 + multipopT (k - 1) s
    + phi (multipop (k - 1) s) - 1 - phi s
-- 1s cancel out
  === multipopT (k - 1) s
    + phi (multipop (k - 1) s) - phi s
-- use induction
  ? multipopP (k - 1) s
  *** QED
\end{lstlisting}

\section{Full proof code of the finger tree}\label{ap:fingertree}

\begin{lstlisting}
data FingerTree a = Empty
  | Single a
  | Deep (Digit a) (FingerTree (Node a)) (Digit a)

data Node a = Node2 a a | Node3 a a a

data Digit a = One a
  | Two a a
  | Three a a a
  | Four a a a a

{-@ reflect cons @-}
cons :: a -> FingerTree a -> FingerTree a
cons a EmptyT = Single a
cons a (Single b) =
  Deep (One a) EmptyT (One b)
cons a (Deep (One b) m sf) =
  Deep (Two a b) m sf
cons a (Deep (Two b c) m sf) =
  Deep (Three a b c) m sf
cons a (Deep (Three b c d) m sf) =
  Deep (Four a b c d) m sf
cons a (Deep (Four b c d e) m sf) =
  Deep (Two a b) (cons (Node3 c d e) m) sf

{-@ reflect snoc @-}
snoc :: FingerTree a -> a -> FingerTree a
snoc EmptyT x = Single x
snoc (Single a) x =
  Deep (One a) EmptyT (One x)
snoc (Deep pr m (One a)) x =
  Deep pr m (Two a x)
snoc (Deep pr m (Two a b)) x =
  Deep pr m (Three a b x)
snoc (Deep pr m (Three a b c)) x =
  Deep pr m (Four a b c x)
snoc (Deep pr m (Four a b c d)) x =
  Deep pr (snoc m (Node3 a b c)) (Two d x)

{-@ reflect dang @-}
{-@ dang :: Digit a -> Nat @-}
dang :: Digit a -> Int
dang One{} = 1
dang Two{} = 0
dang Three{} = 0
dang Four{} = 1

{-@ reflect phi @-}
{-@ phi :: FingerTree a -> Nat @-}
phi :: FingerTree a -> Int
phi EmptyT = 0
phi Single{} = 0
phi (Deep pr m sf) =
  dang pr + phi m + dang sf

{-@ reflect consT @-}
{-@ consT :: a -> FingerTree a -> Nat @-}
consT :: a -> FingerTree a -> Int
consT _ EmptyT = 1
consT _ Single{} = 1
consT _ (Deep One{} _ _) = 1
consT _ (Deep Two{} _ _) = 1
consT _ (Deep Three{} _ _) = 1
consT _ (Deep (Four _ c d e) m _) =
  1 + consT (Node3 c d e) m

{-@ reflect snocT @-}
{-@ snocT :: FingerTree a -> a -> Nat @-}
snocT :: FingerTree a -> a -> Int
snocT EmptyT _ = 1
snocT Single{} _ = 1
snocT (Deep _ _ One{}) _ = 1
snocT (Deep _ _ Two{}) _ = 1
snocT (Deep _ _ Three{}) _ = 1
snocT (Deep _ m (Four a b c _)) _ =
    1 + snocT m (Node3 a b c)

{-@ consP :: x:a -> t:FingerTree a ->
  { consT x t + phi (cons x t) - phi t <= 3 }
@-}
consP :: a -> FingerTree a -> Proof
consP x t@EmptyT =
    consT x t + phi (cons x t) - phi t
-- Use definition of cons/T _ EmptyT
  === 1 + phi (Single x) - phi EmptyT
-- Use definition of phi
  === 1 + 0 - 0 -- trivial, 1 <= 3
  *** QED
consP x t@(Single a) =
    consT x t + phi (cons x t) - phi t
-- Use definition of cons/T _ (Single a)
  === 1 + phi (Deep (One x) EmptyT (One a))
        - phi (Single a)
-- Use definition of phi
  === 1 + dang (One x) + phi EmptyT
        + dang (One a) - 0
-- Use definition of phi and dang
  === 1 + 1 + 0 + 1 - 0 -- trivial, 3 <= 3
  *** QED
consP x t@(Deep (One a) m sf) =
    consT x t + phi (cons x t) - phi t
-- Use definition of cons/T
  === 1 + phi (Deep (Two x a) m sf) - phi t
-- Use definition of phi
  === 1 + dang (Two x a) + phi m + dang sf
    - (dang (One a) + phi m + dang sf)
-- Remove parenthesis, use definition of dang
  === 1 + 0 + phi m + dang sf
    - 1 - phi m - dang sf
-- everything cancels
  === 0 -- trivial, 0 <= 3
  *** QED
consP x t@(Deep (Two a b) m sf) =
    consT x t + phi (cons x t) - phi t
-- Use definition of cons/T
  === 1 + phi (Deep (Three x a b) m sf) - phi t
-- Use definition of phi
  === 1 + dang (Three x a b) + phi m + dang sf
    - (dang (Two a b) + phi m + dang sf)
-- Remove parenthesis, use definition of dang
  === 1 + 0 + phi m + dang sf
    - 0 - phi m - dang sf
-- phi and dang cancels
  === 1 -- trivial, 1 <= 3
  *** QED
consP x t@(Deep (Three a b c) m sf) =
    consT x t + phi (cons x t) - phi t
-- Use definition of cons/T
  === 1 + phi (Deep (Four x a b c) m sf) - phi t
-- Use definition of phi
  === 1 + dang (Four x a b c) + phi m + dang sf
    - (dang (Three a b c) + phi m + dang sf)
-- Remove parenthesis, use definition of dang
  === 1 + 1 + phi m + dang sf
    - 0 - phi m - dang sf
-- phi and dang cancels
  === 2 -- trivial, 2 <= 3
  *** QED
consP x t@(Deep (Four a b c d) m sf) =
    consT x t + phi (cons x t) - phi t
-- Use definition of consT
  === (1 + consT (Node3 b c d) m)
  + phi (cons x t) - phi t
-- Use definition of cons
  === (1 + consT (Node3 b c d) m)
  + phi (Deep (Two x a) (cons (Node3 b c d) m) sf)
  - phi t
-- Use definition of phi
  === (1 + consT (Node3 b c d) m)
  + dang (Two x a) + phi (cons (Node3 b c d) m)
  + dang sf
  - (dang (Four a b c d) + phi m + dang sf)
-- Remove parenthesis
  === 1 + consT (Node3 b c d) m
  + dang (Two x a) + phi (cons (Node3 b c d) m)
  + dang sf
  - dang (Four a b c d) - phi m - dang sf
-- dang sf cancels, use definition of dang
  === 1 + consT (Node3 b c d) m
  + 0 + phi (cons (Node3 b c d) m) - 1 - phi m
-- 1s cancel
  === consT (Node3 b c d) m
  + phi (cons (Node3 b c d) m) - phi m
-- Use induction
  ? consP (Node3 b c d) m
  *** QED

{-@ snocP :: t:FingerTree a -> x:a ->
  { snocT t x + phi (snoc t x) - phi t <= 3 }
@-}
snocP :: a -> FingerTree a -> Proof
snocP t@EmptyT x =
    snocT t x + phi (snoc t x) - phi t
-- Use definition of snoc/T EmptyT _
  === 1 + phi (Single x) - phi EmptyT
-- Use definition of phi
  === 1 + 0 - 0 -- trivial, 1 <= 3
  *** QED
snocP t@(Single a) x =
    snocT t x + phi (snoc t x) - phi t
-- Use definition of snoc/T (Single a) _
  === 1 + phi (Deep (One a) EmptyT (One x))
        - phi (Single a)
-- Use definition of phi
  === 1 + dang (One a) + phi EmptyT
        + dang (One x) - 0
-- Use definition of phi and dang
  === 1 + 1 + 0 + 1 - 0 -- trivial, 3 <= 3
  *** QED
snocP t@(Deep pr m (One a)) x =
    snocT t x + phi (snoc t x) - phi t
-- Use definition of snoc/T
  === 1 + phi (Deep pr m (Two a x)) - phi t
-- Use definition of phi
  === 1 + dang pr + phi m + dang (Two a x)
    - (dang pr + phi m + dang (One a))
-- Remove parenthesis, use definition of dang
  === 1 + dang pr + phi m + 0
    - dang pr - phi m - 1
-- everything cancels
  === 0 -- trivial, 0 <= 3
  *** QED
snocP t@(Deep pr m (Two a b)) x =
    snocT t x + phi (snoc t x) - phi t
-- Use definition of snoc/T
  === 1 + phi (Deep pr m (Three a b x)) - phi t
-- Use definition of phi
  === 1 + dang pr + phi m + dang (Three a b x)
    - (dang pr + phi m + dang (Two a b))
-- Remove parenthesis, use definition of dang
  === 1 + dang pr + phi m + 0
    - dang pr - phi m - 0
-- phi and dang cancels
  === 1 -- trivial, 1 <= 3
  *** QED
snocP t@(Deep pr m (Three a b c)) x =
    snocT t x + phi (snoc t x) - phi t
-- Use definition of snoc/T
  === 1 + phi (Deep pr m (Four a b c x)) - phi t
-- Use definition of phi
  === 1 + dang pr + phi m + dang (Four a b c x)
    - (dang pr + phi m + dang (Three a b c))
-- Remove parenthesis, use definition of dang
  === 1 + dang pr + phi m + 1
    - dang pr - phi m - 0
-- phi and dang cancels
  === 2 -- trivial, 2 <= 3
  *** QED
snocP t@(Deep pr m (Four a b c d)) x =
    snocT t x + phi (snoc t x) - phi t
-- Use definition of snocT
  === (1 + snocT m (Node3 a b c))
  + phi (snoc t x) - phi t
-- Use definition of snoc
  === (1 + snocT m (Node3 a b c))
  + phi (Deep pr (snoc m (Node3 a b c)) (Two x a))
  - phi t
-- Use definition of phi
  === (1 + snocT m (Node3 a b c))
  + dang pr + phi (snoc m (Node3 a b c))
  + dang (Two x a)
  - (dang pr + phi m + dang (Four a b c d))
-- Remove parenthesis
  === 1 + snocT m (Node3 a b c)
  + dang pr + phi (snoc m (Node3 a b c))
  + dang (Two x a)
  - dang pr - phi m - dang (Four a b c d))
-- dang pr cancels, use definition of dang
  === 1 + snocT m (Node3 a b c)
  + phi (snoc m (Node3 a b c)) + 0 - phi m - 1
-- 1s cancel
  === snocT m (Node3 a b c)
  + phi (snoc m (Node3 a b c)) - phi m
-- Use induction
  ? snocP m (Node3 a b c)
  *** QED

\end{lstlisting}

\end{document}
\endinput